\documentclass[paper=a4, fontsize=11pt]{jhwhw} % A4 paper and 11pt font size
\usepackage{amsmath,amsfonts,amsthm, amssymb} % Math packages
\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{enumerate}
\usepackage{mathtools}
\usepackage{color}
\newcommand\SetSymbol[1][]{\:#1\vert\:}
\providecommand\given{} % to make it exist
\DeclarePairedDelimiterX\Set[1]\{\}{\renewcommand\given{\SetSymbol[\delimsize]}#1}

\begin{document}
\title{Real Analysis - Assignment \# 3}
\author{Ben Haines}

\problem{1}
Let $A$ and $B$ be non-empty subsets of $\mathbb R$ such that $a\le b$ for all $a\in A$ and $b\in B$. Prove that $A$ is bounded above, $B$ is bounded below and $\sup(A)\le \inf(B)$. \textbf{Hint: }This can be proved directly from the definitions of supremum and infimum without any computations or using any theorems, but you need to proceed in two steps.
\solution
Select any element $b\in B$. Then for all $x\in A$, $x\le b$ so $b$ is an upper bound for $A$. Now select any element $a\in A$. Then for all $x\in B$, $a \le x$ so $a$ is a lower bound for $B$. Assume that $\inf(B) < \sup(A)$. Then there must exist some $x\in A$ such that $x> \inf(B)$ otherwise there would be an upper bound for $A$ less than $\sup(A)$. However, this means that $x$ is a lower bound for $B$ and $x > \inf(B)$. This is not possible. Thus $\sup(A) \le \inf(B)$. 

\problem{2}
Use induction to prove the formlua for the sum of a (finite) geometric progression: $a + ar + ar^2 + \ldots + ar^{n-1} = a\frac{1-r^n}{1-r}$ where $a,r \in \mathbb R$ and $r\not= 1$.
\solution
Base Case P(1):
$$a = a\frac{1-r}{1-r} = a$$
Assume P(k):
$$a + ar + \ldots + ar^{k-1} = a\frac{1-r^k}{1-r}$$
Prove P(k+1):
\begin{align*}
    a + ar + \ldots + ar^{k-1} + ar^k &= a\frac{1-r^k}{1-r} + ar^k\\
                                      &= a\left(\frac{1-r^k}{1-r} + r^k\right)\\
                                      &= a\left(\frac{1-r^k}{1-r} + \frac{(1-r)r^k}{1-r}\right)\\
                                      &= a\left(\frac{1-r^{k+1}}{1-r}\right)
\end{align*}


\problem{3}
Prove the following inequalities by induction:
\begin{enumerate}[(i)]
    \item $n < 2^n$ for all $n\in \mathbb N$
    \item $n^2 < 2^n$ for all integers $n\ge 4$ 
\end{enumerate}
\solution
\begin{enumerate}[i]
    \item 
        Base Case:
        $$1 < 2$$
        Assume P(k):
        $$k < 2^k$$
        Prove P(k+1):
        $$k < 2^k \implies 2k < 2^{k+1}$$
        $k\ge 1$ so $k+1 \le 2k$. Then by the transitive property, $k+1 < 2^{k+1}$.
    \item 
        Something wrong with the problem.
\end{enumerate}


\problem{4}
Use induction to prove Bernoulli's inequality:
$$(1 + x)^n \ge 1 + nx\text{ for all }n\in \mathbb N\text{ and } x\ge -1$$
\solution
Base Case:
$$1 = (1+x)^0 \ge 1 = 1 + 0x$$
Assume P(k):
$$(1+x)^k \ge 1 + kx$$
Prove P(k+1):
\begin{align}
    (1 + x)^{k+1} &= (1+x)^k(1+x)\\
                  &\ge (1 + kx)(1+x)\\
    (1 + kx)(1+x) &= 1 + kx + kx^2\\
                  &= 1 + (k+1)x + kx^2\\
                  &\ge 1 + (k+1)x
\end{align}
So $(1+x)^{k+1} \ge 1 + (k + 1)x$

\problem{5}
Prove that the sequence converges to $L$, and explicitly find a function $M(\epsilon)$ satisfying $(1')$ above.
\begin{enumerate}[(i)]
    \item $a_n = \frac{2n^2 + 3}{n^2 - n - \cos(n)}$, $L = 2$
    \item $a_n = \frac{n}{4^n}$, $L = 0$
\end{enumerate}
\solution
\begin{enumerate}[(i)]
    \item Fix $\epsilon > 0$. $|a_n - 2| = \left|\frac{2n^2 + 3}{n^2 - n - \cos(n)} - 2\right| = \left|\frac{2n + 2\cos(n) + 3}{n^2-n-\cos(n)}\right|$. We can bound this fraction above by $\frac{7n}{n^2}$. Then we want $n$ such that $\frac{7}{n} < \epsilon$. 

            Let $M(\epsilon) = \frac{7}{\epsilon}$ then $\forall n > M(\epsilon)$ we have $n > \frac{7}{\epsilon} \implies \epsilon > \frac{7}{n}$. Then $|a_n - 2| < \frac{7}{n} < \epsilon$.

        \item $\left|\frac{n}{n^4}\right|$ can be bounded above by $\left|\frac{n^3}{n^4}\right| = \frac{1}{n}$ Let $M(\epsilon) = \frac{1}{\epsilon}$. Then $\forall n > M(\epsilon)$, $n > \frac{1}{\epsilon} \implies \epsilon > \frac{1}{n}$. Then from above $|a_n - 0| \le \frac{1}{n} < \epsilon$.
\end{enumerate}

\problem{6}
Let ${a_n}$ and ${b_n}$ be sequences. Suppose that for every $\epsilon > 0$ the following is true: $|a_n - 3| < \epsilon$ for all $n > \frac{10}{e^2}$ and $|b_n - 4| < \epsilon$ for all $n> \frac{1}{e^3}$. Find an explicit function $M(\epsilon)$ such that $|a_n + b_n - 7| < \epsilon$ for all $n > M(\epsilon)$. 
\solution
Let $\epsilon' = \epsilon/2$. This is guaranteed to be a number larger than 0. Then $|a_n - 3| < \epsilon'$ for all $n > (10/(e/2)^2)$ and $|b_n - 4| < \epsilon'$ for all $n > (1/(\epsilon/2)^3)$. Then by Theorem 2.12 $|a_b + b_n - 7| < \epsilon' + \epsilon' = \epsilon$ for $n > \max((10/(e/2)^2), (1/(\epsilon/2)^3))$. Define $M(\epsilon) = \max((10/(e/2)^2), (1/(\epsilon/2)^3))$.

\problem{7}
Let ${a_n}$ be a sequence, and define $b_k$ and $c_k$ (with $k\in \mathbb N$) by $b_k = a_{2k-1}$ and $c_k = a_{2k}$, that is ${b_k}$ and ${c_k}$ are subsequences of ${a_n}$ consisting of its elements located in odd (respectively even) position. Suppose that ${b_k}$ and ${c_k}$ both converge and $\lim_{k\to \infty}b_k = \lim_{k\to \infty}c_k = L$ for some $L\in \mathbb R$. Prove that ${a_n}$ converges to $L$ as well. 
\solution
For all $\epsilon > 0$ there exist $R_1, R_2$ such that $|b_k - L| < \epsilon$ for all $n > R_1$ and $|c_k - L| < \epsilon$ for all $n > R_2$. Let $M = \max(R_1, R_2)$. Then for any $a_n$ where $n > M$ if $a_n\in \{b_k\}$ then $|a_n - L| < \epsilon$. Similarly if $a_n\in \{c_k\}$ then $|a_n - L| < \epsilon$. Therefore the limit of $\{a_n\}$ is $L$.

\problem{8}
Let $f: X\to Y$ be a function. Prove that the following conditions are equivalent:
\begin{enumerate}[(a)]
    \item $f$ is injective
    \item $f(A\cap C) = f(A)\cap f(C)$ for any two subsets $A, C$ of $X$.
\end{enumerate}
\solution
First let $f$ be injective and consider any element $x\in f(A\cap C)$. $x = f(y)$ for some $y\in A\cap C$. Then $y\in A\implies x\in f(A)$ and $y\in B\implies x\in f(C)$ so $x\in f(A)\cap f(C)$. Now consider any $x\in f(A)\cap f(C)$. Then $x = f(a)$ for some $a\in A$ and $x = f(c)$ for some $c\in C$. $f$ is injective so $a = c\in A$ and $a=c\in C$. Then $x\in f(A\cap C)$. 

The second part is proved by contraposition. Assume that $f$ is not injective. Then we must show there are two subsets $A$ and $C$ of $X$ such that it is not the case that $f(A\cap C) = f(A) \cap f(C)$. We know by the fact that $f$ is not injective that there exist two elements $x_1, x_2\in X$ such that $f(x_1) = f(x_2)$ but $x_1 != x_2$. Let $A=\Set{x_1}$ and $C = \Set{x_2}$. Then $f(A\cap C) = \emptyset \not= f(A) \cap f(C) = \Set{f(x_1)}$. 
\end{document}
