\documentclass{article}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{color}
\usepackage{listings}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=C++,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  numbers=left,
  tabsize=3
}

\pagestyle{fancy}
\fancyhf{}
\chead{inlab9.pdf}
\lhead{Ben Haines, bmh5wx}
\rhead{11/4/14}
\begin{document}
\section{Inheritance}
Consider the following example of simple inheritance in c++. It includes a derived class "Child" that inherits from the base class "Parent." It inherits Parent's x and y members and also has a public member of its own, "luckyFin."
\begin{lstlisting}
class Parent 
{
   public:
      void setX(int w)
      {
         x = w;
      }
      void setY(int h)
      {
         y = h;
      }
   protected:
      int x;
      int y;
};

// Derived class
class Child: public Parent
{
   public:
      int xPlusY()
      { 
         return (x + y); 
      }

      int luckyFin;
};

int main(void)
{
   Child * nemo = new Child();
 
   nemo->setX(2);
   nemo->setY(3);

   int add = nemo->xPlusY();

   return 0;
}
\end{lstlisting}
\lstset{language={[x86masm]Assembler}}
The resulting assembly code is too long to list in full but snippets will be shown as the topics they are relevant to are discussed.
\subsection{Initialization}
The following x86 snippet shows the initializtion of the nemo object.
\begin{lstlisting}
	call	_Znwj
	add	esp, 16
	mov	DWORD PTR [eax], 0
	mov	DWORD PTR [eax+4], 0
	mov	DWORD PTR [eax+8], 0
	mov	DWORD PTR [ebp-12], eax
	mov	eax, DWORD PTR [ebp-12]
\end{lstlisting}
The call to _Znwj returns a pointer to the location of the newly created nemo object in memory. Then the data members of nemo are moved into memory at that location. 

\subsection{Data Layout}
From the code above it is not immediately clear which element is stored at which location. This information can be deduced by considering a call to setX inherited from the Parent class. The assembly for such a call is as follows.
\begin{lstlisting}
	push	2
	push	eax
	call	_ZN6Parent4setXEi
\end{lstlisting}
with the function itself being
\begin{lstlisting}
_ZN6Parent4setXEi:
	push	ebp
	mov	ebp, esp
	mov	eax, DWORD PTR [ebp+8]
	mov	edx, DWORD PTR [ebp+12]
	mov	DWORD PTR [eax], edx
	pop	ebp
	ret
\end{lstlisting}
Before the call is made, eax holds the address of the nemo object in memory. The setX method then moves the argument '2' located at [ebp+12] into [eax], the lowest memory location in nemo. Similar analysis of the setY function shows us that $y$ is located at [eax+4]. From this we can assume that 'luckyFin' is stored at [eax+8]. 

The general rule that we can conclude from this examination is that within an object that inherits from a parent class the object's own member variables are stored first (at higher memory addresses) and then the inherited members are stored in reverse order with the first member of the parent class being stored at the lowest address.

\subsection{Destruction}
The first example will demonstrate what happens when a user created object goes out of scope. In order to more easily demonstrate this the c++ code has been modified in the following ways:

\begin{lstlisting}
void scope() {
    Child * nemo = new Child();
    nemo->setX(2);
    nemo->setY(3);
    int add = nemo->xPlusY();
}

int main(void) {
    scope();
    return 0;
}
\end{lstlisting}
The last lines of assembly within the 'scope' function call look like this:
\begin{lstlisting}
	call	_ZN5Child6xPlusYEv
	add	esp, 16
	mov	DWORD PTR [ebp-16], eax
	leave
	ret
\end{lstlisting} 
As can be seen, the deallocation of the nemo object is very simple and operates just like deallocation of any other variables on the stack by adding to esp.

The second example will show what happens when something is 'destroyed' by the destructor.
\begin{lstlisting}
class Planet {
   public:
      void setX(int w) {
         x = w;
      }
      void setY(int h) {
         y = h;
      }
   protected:
      int x;
      int y;
};

// Derived class
class DevourerOfWorlds {
   public:
      int luckyFin;
      DevourerOfWorlds();
      ~DevourerOfWorlds();

    private:
      Planet * earth;
};

DevourerOfWorlds::DevourerOfWorlds() {
    earth = new Planet();
}

DevourerOfWorlds::~DevourerOfWorlds() {
    delete earth;
}

void scope() {
    DevourerOfWorlds galactus = DevourerOfWorlds();
}

int main(void) {
    scope();
    return 0;
}

\end{lstlisting}
Here the constructors and destructors have been explicitly defined. The destructor is automatically called when the program exits the scope of scope(). They compile to
\begin{lstlisting}
_ZN16DevourerOfWorldsC2Ev:
	push	ebp
	mov	ebp, esp
	sub	esp, 8
	sub	esp, 12
	push	8
	call	_Znwj
	add	esp, 16
	mov	DWORD PTR [eax], 0
	mov	DWORD PTR [eax+4], 0
	mov	edx, DWORD PTR [ebp+8]
	mov	DWORD PTR [edx+4], eax
	leave
	ret
_ZN16DevourerOfWorldsD2Ev:
	push	ebp
	mov	ebp, esp
	sub	esp, 8
	mov	eax, DWORD PTR [ebp+8]
	mov	eax, DWORD PTR [eax+4]
	sub	esp, 12
	push	eax
	call	_ZdlPv
	add	esp, 16
	leave
	ret
\end{lstlisting}
The constructor initializes galactus as we expect and also calls _Znwj to create the Planet earth object. 

In the destructor, the address of earth is moved into eax which is then passed as an argument to _ZdlPv which deallocates the memory allocated to it. From there things proceed as normal with memory being deallocated by adding to esp and then leave and ret making up the standard epilogue.
\end{document}
